### 2013-03-26 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
starting work on parallel rngs but I need a few references for this

%%---------------------------------------------------------------------------%%
\section{Fully Asynchronous MSOD Neumann-Ulam Algorithm}
\label{sec:asynchronous_algorithm}
We can take much of what was learned from Brunner and Wagner's
parallel Monte Carlo methods for radiation transport and directly
apply it to a parallel formulation of our stochastic linear
solvers. Direct analogs can be derived from these works by noting that
the primary difference between solving a linear system with Monte
Carlo methods and fixed source Monte Carlo transport problems is the
content of the Markov chains that are generated. The transitions
represented by these chains are bound by probabilities and weights and
are initiated by the sampling of a source. In the context of transport
problems, those transitions represent events such as particle
scattering and absorption with probabilities that are determined by
physical data in the form of cross sections. For stochastic matrix
inversion, those transitions represent moving between the equations of
the linear system (and therefore the physical domain which they
represent) and their probabilities are defined by the coefficients of
those equations. Ultimately, we tally the contributions to generate
expectation values in the desired states as we progress through the
chains. Therefore, parallel methods for Monte Carlo radiation
transport can be abstracted and we can use those concepts that apply
to matrix inversion methods as an initial means of developing a
parallel Neumann-Ulam-type solver.

\subsection{Parallel Random Number Generation}
\label{subsec:parallel_rng}

@@

%%---------------------------------------------------------------------------%%
\section{Parallel MCSA}
\label{sec:parallel_mcsa}
With a parallel adjoint Neumann-Ulam solver implementation, the
parallel implementation of the MCSA method will be trivial. Recall the
MCSA iteration procedure outlined in Eq~(\ref{eq:mcsa}). In
\S~\ref{sec:parallel_krylov_methods} we discussed parallel matrix and
vector operations as utilized in conventional Krylov methods. We
utilize these here for the parallel MCSA implementation. In the first
step, a parallel matrix-vector multiply is used to apply the split
operator to the previous iterate's solution. A parallel vector update
is then performed with the source vector to arrive at the initial
iteration guess. In the next step, the residual is computed by the
same operations where now the operator is applied to the solution
guess with a parallel matrix-vector multiply and then a parallel
vector update with the source vector is performed. Once the correction
is computed with a parallel adjoint Neumann-Ulam solve, this
correction is applied to the guess with a parallel vector update to
get the new iteration solution. Additionally, as given by
Eq~(\ref{eq:mcsa_stopping_criteria}), 2 parallel vector reductions
will be required to check the stopping criteria: one initially to
compute the infinity norm of the source vector, and another at every
iteration to compute the infinity norm of the residual vector. For
this implementation, all of the issues that will be potentially
generated by the parallel adjoint solver implementation will manifest
themselves here as the quality of the correction will be of intense
study.

In addition to parallel implementation and performance, MCSA's
potential for aiding advancement in non-symmetric matrix solutions
leads to a natural comparison with the GMRES algorithm. As both
solvers are aimed at the same class of problem, we desire a set of
metrics that will allow us to quantitatively compare the two. Given
that the Krylov subspace maintained by GMRES can become large, do we
benefit from a memory standpoint with an MCSA scheme in that no
subspace is required? Does this benefit outweigh the fact that the
linear operator must be explicitly formed in order to build the
transition probabilities for the random walk sequence? For
non-symmetric systems, does MCSA exhibit similar convergence
properties to Krylov methods? If a Krylov methods build a subspace,
can those memory savings in MCSA be used to implement domain
replication in the adjoint solver? Such questions can be answered by a
comparative study of the two solvers that controls the system size and
the iterations required to converge.

>>>>

%%---------------------------------------------------------------------------%%
\section{Fully Asynchronous MSOD Neumann-Ulam Algorithm}
\label{sec:asynchronous_algorithm}
We can take much of what was learned from Brunner and Wagner's
parallel Monte Carlo methods for radiation transport and directly
apply it to a parallel formulation of our stochastic linear
solvers. Direct analogs can be derived from these works by noting that
the primary difference between solving a linear system with Monte
Carlo methods and fixed source Monte Carlo transport problems is the
content of the Markov chains that are generated. The transitions
represented by these chains are bound by probabilities and weights and
are initiated by the sampling of a source. In the context of transport
problems, those transitions represent events such as particle
scattering and absorption with probabilities that are determined by
physical data in the form of cross sections. For stochastic matrix
inversion, those transitions represent moving between the equations of
the linear system (and therefore the physical domain which they
represent) and their probabilities are defined by the coefficients of
those equations. Ultimately, we tally the contributions to generate
expectation values in the desired states as we progress through the
chains. Therefore, parallel methods for Monte Carlo radiation
transport can be abstracted and we can use those concepts that apply
to matrix inversion methods as an initial means of developing a
parallel Neumann-Ulam-type solver. In this section the fully
asynchronous MSOD Neumann-Ulam algorithm developed for this work is
presented step-by-step in detail.

\subsection{Parallel Random Number Generation}
\label{subsec:parallel_rng}
Correct parallel Monte Carlo requires parallel random number
generation. To ensure distinct Markov chains for each history in the
problem, regardless of its state in phase space or parallel space, a
unique stream of random numbers is necessary to achieve uncorrelated
tally results for the solution.

@@

%%---------------------------------------------------------------------------%%
\section{Parallel MCSA}
\label{sec:parallel_mcsa}
With the parallel adjoint Neumann-Ulam solver implementation described
aboive, the parallel implementation of the MCSA method is
trivial. Recall the MCSA iteration procedure outlined in
Eq~(\ref{eq:mcsa}). In \S\ref{sec:parallel_krylov_methods} we
discussed parallel matrix and vector operations as utilized in
conventional Krylov methods. We utilize these here for the parallel
MCSA implementation. In the first step, a parallel matrix-vector
multiply is used to apply the split operator to the previous iterate's
solution. A parallel vector update is then performed with the source
vector to arrive at the initial iteration guess. In the next step, the
residual is computed by the same operations where now the operator is
applied to the solution guess with a parallel matrix-vector multiply
and then a parallel vector update with the source vector is
performed. Once the correction is computed with a parallel adjoint
Neumann-Ulam solve, this correction is applied to the guess with a
parallel vector update to get the new iteration
solution. Additionally, as given by
Eq~(\ref{eq:mcsa_stopping_criteria}), 2 parallel vector reductions
will be required to check the stopping criteria: one initially to
compute the infinity norm of the source vector, and another at every
iteration to compute the infinity norm of the residual vector. For
this implementation, all of the issues that will be potentially
generated by the parallel adjoint solver implementation will manifest
themselves here as the quality of the correction will be of intense
study.
